Matching Bullets

### 2015-12-17 imaging-paper.Rnw
probability ratios

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 36 pairs of lands with at least sixteen CMS in common are matches. There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches. On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}). Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}. Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}. Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae. Visually we can easily tell these two lands do not match well.
%
\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{Distribution of maximal CMS.}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer(palette="Paired") +
  xlab("maximum CMS")
@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

@@

For smaller numbers of CMS, the percentage of false positives quickly increases. However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two cross sections (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:distr.dist}Relationship between maximal CMS, distances between cross sections and matches.}
<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match), alpha=0.5) +
   facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
  ylab("Distance between cross sections") +
  scale_colour_brewer(palette="Paired") +
  theme_bw() + 
  theme(legend.position="none") +
    xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:ccfs-match}Relationship between maximal CMS, cross-correlation between cross sections and matches.}
<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + theme_bw() + 
  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + 
  facet_wrap(~match, labeller="label_both") + 
    scale_colour_brewer(palette="Paired") +
  xlab("maximum CMS") + ylab("Cross-correlation") +
  theme(legend.position="none")
@
\end{subfigure}
\caption{A small distance between cross sections or a large cross-correlation between cross sections are indicative of matches between lands.}
\end{figure}

>>>>

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 36 pairs of lands with at least sixteen CMS in common are matches. There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches. On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}). Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}. Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}. Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae. Visually we can easily tell these two lands do not match well.
%
\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{Distribution of maximal CMS.}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer(palette="Paired") +
  xlab("maximum CMS")

@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=cms, fill=match), alpha=0.5) +
  xlab("maximum CMS") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

@@

For smaller numbers of CMS, the percentage of false positives quickly increases. However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two cross sections (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match), alpha=0.5) +
   facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
  ylab("Distance between cross sections") +
  scale_colour_brewer(palette="Paired") +
  theme_bw() + 
  theme(legend.position="none") +
    xlab("maximum CMS")

@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=distr.dist, fill=match), alpha=0.5) +
  xlab("Distances between cross sections") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
\end{figure}

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + theme_bw() + 
  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + 
  facet_wrap(~match, labeller="label_both") + 
    scale_colour_brewer(palette="Paired") +
  xlab("maximum CMS") + ylab("Cross-correlation") +
  theme(legend.position="none")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=ccf, fill=match), alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
x <- seq(0,1, by=0.001)
library(sm)
km <- sm.density(bstats$ccf[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$ccf[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

densities <- data.frame(
  x = rep(c(0, x, 1), times = 2),
  estimate = c(0, km$estimate, 0, 0, knm$estimate,0),
  match = rep(c("KM", "KNM"), each=1003))

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match, colour=match), 
               data=densities, alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  ylab("Density estimate") +
  theme(legend.position=c(.975,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x, pmin(5, kmtail/knmtail)), data=dframe) 
@
\end{subfigure}
\caption{A large cross-correlation between cross sections is indicative of matches between lands.}
\end{figure}
